Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-03-15 main.tex
In the middle of working on changing terminology

I summary, the alignment stage essentially two levels of available parallelism:
at the higher level there are multiple alignent problems that are solved
independently.  At a lower level, the first-order linear algebraic operations
exhibit parallelism within image operations, i.e. at a pixel level.

@@

This section discusses how the iterative alignment problem can be implemented
on CPU hardware.  The implementations presented in \cite{WagnerA2009-CVPR} and
\cite{WagnerA2011-PAMI} were written using the OpenCV library.

OpenCV wraps either the operating system's built-in BLAS library, or Math
Kernel Library (MKL), Intel's BLAS implementation, and thus achieves vector
(i.e. pixel)-level parallelism implicitly.  The authors of
\cite{WagnerA2011-PAMI} used OpenMP to schedule instances of the alignment
problem to the CPU cores in a 1:1 fashion.  With eight alignment problems for
ten training users being solved concurrently, and with 20-40 images per
training user, all of the required data fits into L3 cache.  This provided a
significant, though sub-linear improvement over a single-core implementation.

@@

In the proposed CPU implementation of the iterative alignment stage, we
therefore perform parallelization in a manner similar to the matlab code:
solving a single problem instance at a time, relying on the threaded version of
MKL BLAS for both multi-core and vector parallelism.  Why is this
implementation faster?  Recall that the L2 cache can old up to 128 image
vectors' worth of data.  Since we're only aligning data for 1 gallery user at a
time the working set of the problem fits into L2 cache, resulting in
significantly higher image alignment speed than previous cpu implementations.

@@

In this section, we benchmark the performance of our algorithm on CPU and GPU
platforms.  In order to show how our $\ell_1$ minimization algorithms scale
with problem size, in Section \ref{sec:simulation} we will begin with
benchmarks on synthetic data.  We will then progress in Sections
\ref{sec:alignment_benchmark} and \ref{sec:recognition_benchmark} to
demonstrating the speed of our implementations as applied to the alignment and
recognition stages of the face recognition problem.

>>>>

I summary, the alignment stage essentially two levels of available parallelism:
at the higher level there are multiple alignent problems that are solved
independently, or {\bf problem-level} parallelism.  At a lower level, the first-order linear algebraic operations
exhibit parallelism within image operations, i.e. at a pixel level.  We term
this {\bf pixel-level} parallelism.

@@

One of the main contributions of this paper is a more efficent mapping of the
two levels of parallelism available in iterative face alignment to the two
levels of concurrency available in atypical multicore CPU architecture.

We argue that, in constrast with previous implementations, which mapped problem
level parallelism onto core-level concurrency, it is more efficient to
completely disregard problem-level parallelism, and instead map pixel-level
parallelism onto both core-level and vector-level cpu concurrency.  In other words,
the implementation we advocate performs alignment of the test image against the gallery
images of a single user at a time.  Each matrix-vector and vector-vector operation
gets broken into 8 parts that run one per core, and into 4 single float vectors
that execute on the SSE SIMD units.

Why is this
implementation faster?  Recall that the L2 cache can old up to 128 image
vectors' worth of data.  Since we're only aligning data for one gallery user at a
time the working set of the problem fits into L2 cache, resulting in
significantly higher image alignment speed than previous cpu implementations.

In our proposed CPU implementation of the iterative alignment stage, we
therefore perform parallelization in a manner similar to the matlab code:
solving a single problem instance at a time, relying on the threaded version of
MKL BLAS for both multi-core and vector parallelism.

This section discusses how the iterative alignment problem can be implemented
on CPU hardware.

{\bf Comparison to Previous Implementations}
The implementations presented in \cite{WagnerA2009-CVPR} and
\cite{WagnerA2011-PAMI} were written using the OpenCV library.  OpenCV wraps
either the operating system's built-in BLAS library, or Math Kernel Library
(MKL), Intel's BLAS implementation, and thus achieves vector (i.e. pixel)-level
parallelism implicitly.  The authors of \cite{WagnerA2011-PAMI} used OpenMP to
schedule instances of the alignment problem to the CPU cores in a 1:1 fashion.
With eight alignment problems for ten training users being solved concurrently,
and with 20-40 images per training user, all of the required data fits into L3
cache.  This provided a significant, though sub-linear improvement over a
single-core implementation.

@@

@@

In this section, we benchmark the performance of our parallel implementations
of $\ell_1$ minimization on CPU and GPU platforms.  In order to show how our
$\ell_1$ minimization algorithms scale with problem size, in Section
\ref{sec:simulation} we will begin with benchmarks on synthetic data.  We will
then progress in Sections \ref{sec:alignment_benchmark} and
\ref{sec:recognition_benchmark} to demonstrating the speed of our
implementations as applied to the alignment and recognition stages of the face
recognition problem.
